LAGOS lake depth

### 2019-10-04 lagos_depth.Rmd
viz hypso collection

```{r timings, include=FALSE, eval=FALSE}
# https://stackoverflow.com/a/24600461/3362993
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste('Time for this code chunk:', as.character(round(res, 2)))
    }
  }})
)
```

```{r setup, include=FALSE}
source("scripts/99_utils.R")
knitr::opts_chunk$set(echo = FALSE, timeit = TRUE, eval = FALSE)
```

```{r data, message=FALSE, echo=FALSE, eval=TRUE}
dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  mutate(max_depth_m_class = cut(max_depth_m, 
                     round(quantile(max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
                                    na.rm = TRUE)))) %>% 
    mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 400,
                                                          1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>%
  coordinatize(latname = "lat", longname = "long")
```

@@

```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, eval=TRUE}
knitr::kable(t(quantile(dt$max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
                        na.rm = TRUE)), caption = "Lake depth quantiles")
knitr::kable(round(t(quantile(dt$lake_waterarea_ha, c(0, 0.05, 0.5, 0.95, 1), 
                        na.rm = TRUE)), 2), caption = "Lake area quantiles")
```

@@

ggplot() + 
  geom_sf(data = dt_plot, aes(fill = n_class)) +
  labs(fill = "") +
  ggtitle("Number of lakes with max depth")

dt_plot <- dt %>%
  st_drop_geometry() %>%
  dplyr::filter(!is.na(mean_depth_m)) %>%
  group_by(state) %>%
  tally() %>%
  mutate(n_class = smart_cut(.$n, 5, "groups", 
                        labels = ~paste(sep = "-", round(.y[1]), round(.y[2])))) %>%
  right_join(usa_sf()) %>%
  st_set_geometry("geometry")

@@

```{r cutoffs, eval = TRUE, echo=FALSE}

@@

>>>>

```{r setup, include=FALSE}
source("scripts/99_utils.R")
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, message = FALSE,
                      fig.path = "figures/", fig.width = 5, fig.height = 6,
                      dev = c("pdf"))
```

```{r data, message=FALSE, echo=FALSE, eval=TRUE}
dt <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  mutate(max_depth_m_class = cut(max_depth_m, 
                     round(quantile(max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
                                    na.rm = TRUE)))) %>% 
    mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 400,
                                                          1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>%
  coordinatize(latname = "lat", longname = "long")

hypso_ct <- read.csv("data/ct_hypso.csv", stringsAsFactors = FALSE)
hypso_mn <- read.csv("data/mn_hypso.csv", stringsAsFactors = FALSE)
```

@@

```{r viz, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, eval=FALSE}
knitr::kable(t(quantile(dt$max_depth_m, c(0, 0.05, 0.5, 0.95, 1), 
                        na.rm = TRUE)), caption = "Lake depth quantiles")
knitr::kable(round(t(quantile(dt$lake_waterarea_ha, c(0, 0.05, 0.5, 0.95, 1), 
                        na.rm = TRUE)), 2), caption = "Lake area quantiles")
```

@@

ggplot() + 
  geom_sf(data = dt_plot, aes(fill = n_class)) +
  labs(fill = "") +
  ggtitle("Number of lakes with max depth")
```

```{r num-lakes-mean, eval=TRUE}
dt_plot <- dt %>%
  st_drop_geometry() %>%
  dplyr::filter(!is.na(mean_depth_m)) %>%
  group_by(state) %>%
  tally() %>%
  mutate(n_class = smart_cut(.$n, 5, "groups", 
                        labels = ~paste(sep = "-", round(.y[1]), round(.y[2])))) %>%
  right_join(usa_sf()) %>%
  st_set_geometry("geometry")

@@

```{r cutoffs, eval = TRUE, echo=FALSE, fig.height = 4}

@@

```{r hypsography, eval=TRUE, fig.height = 4}
ids <- factor(c("concave", "convex"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 75, 0, 
        25, 100, 100), 
  y = c(75, 0, 0, 
        100, 100, 25))

plot_grid(
ggplot(data = hypso_ct) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = id)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("CT", " (", length(unique(hypso_ct$id)), ") lakes")) + 
  theme(legend.position = "bottom"), 
ggplot(data = hypso_mn) +
    geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  ylab("Depth (percent)") + xlab("Area (percent)") +
  ggtitle(paste0("MN", " (", length(unique(hypso_mn$llid)), ") lakes")) +
  theme(axis.title.y = element_blank()), 
rel_widths = c(1, 0.7))
